---
title: "ggplot 1"
output: github_document
---

```{r}
library(tidyverse)
```


load in a dataset
```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USC00519397", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(
      id, 
      USW00094728 = "CentralPark_NY", 
      USC00519397 = "Waikiki_HA",
      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())
```



## Scatterplot

tmax vs. tmin

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point()
```

save gglots

```{r}
ggp_tmax_tmin = 
  weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point()

ggp_tmax_tmin

```


## fancy it up

add lines, color...
```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) + 
  geom_point(alpha = .3) +
  geom_smooth(se = FALSE) + 
  facet_grid(. ~ name)
```

if you put `color` at geom_point, means you only apply it in **point**, while if you put `color` at gglot, it means the color will be applied into every function below. 
in this case, the `geom_smooth` didn't differentiate 3 names by color actually.
```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name), alpha = .3) +
  geom_smooth(se = FALSE)
```


make one more scatterplot

```{r}
weather_df %>% 
  ggplot(aes(x = date, y = tmax, size = prcp)) +
  geom_point(alpha = .3) + 
  facet_grid(. ~ name) +
  geom_smooth(se = FALSE)

```

## use data manipulation as part of this

```{r}
weather_df %>% 
  filter(name == "CentralPark_NY") %>% 
  mutate(
    tmax = tmax * (9 / 5) +32,
    tmin = tmin * (9 / 5) +32
  ) %>% 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point()
```

## stacking geoms

```{r}
weather_df %>% 
  ggplot(aes(x = date, y = tmax, color = name)) + 
  geom_smooth()
```

```{r}
ggplot(weather_df, aes(x = tmax, y = tmin)) + 
  geom_hex()

ggplot(weather_df, aes(x = tmax, y = tmin)) + 
  geom_bin_2d()
```

## Univariate plots

```{r}
weather_df %>% 
  ggplot(aes(x = tmax, fill = name)) + 
  geom_histogram() +
  facet_grid(. ~ name)
```

try some other plots
```{r}
weather_df %>% 
  ggplot(aes(x = tmax, fill = name)) + 
  geom_density(alpha = .3)
```


still with `tmax` and `name`

```{r}
weather_df %>% 
  ggplot(aes(x = name, y = tmax)) + 
  geom_boxplot()
```

some people like violin plots

```{r}
weather_df %>% 
  ggplot(aes(x = name, y = tmax)) + 
  geom_violin()
```

what about ridges ...
```{r}
library(ggridges)
weather_df %>% 
  ggplot(aes(x = tmax, y = name)) +
  geom_density_ridges(alpha = .8, scale = .8)
```

## Embeding plots
1) chunk option
picture become wider, while points become smaller.
```{r, fig.width = 12}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) + 
  geom_point(alpha = .3)

```

maybe try this?

```{r}
knitr::opts_chunk$set(
fig.width = .6,
fig.asp = .6,
out.width = "90%"
)
```
