---
title: "ggplot 1"
output: github_document
---

```{r}
library(tidyverse)
```


load in a dataset
```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USC00519397", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(
      id, 
      USW00094728 = "CentralPark_NY", 
      USC00519397 = "Waikiki_HA",
      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())
```

## start with a familiar one

**scale_x_continuous**

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = .3) +
  labs(
    title = "Temperature at three stations",
    x = "Mininmum daily temp (C)",
    y = "Maximum daily temp (C)",
    caption = "Data from rnoaa package with three stations"
  ) +
  scale_x_continuous(  
    breaks = c(-15, 0 ,15),
    labels = c("-15 C", "0", "15C")  ## Position scales for continuous data (x & y)
  ) +
  scale_y_continuous(
    trans = "sqrt",
    position = "right"  ## put the y axis at the right of the picture
  )

```

Color scales  **scale_color_hue**

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = .3) +
  labs(
    title = "Temperature at three stations",
    x = "Mininmum daily temp (C)",
    y = "Maximum daily temp (C)",
    caption = "Data from rnoaa package with three stations"
  ) + 
  scale_color_hue(
    name = "location",  ## you can change the name above `legend` using this function
    h = c(100, 300)
    )
```

Better using package **viridis**!
```{r}
library(viridis)
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = .3) +
  labs(
    title = "Temperature at three stations",
    x = "Mininmum daily temp (C)",
    y = "Maximum daily temp (C)",
    caption = "Data from rnoaa package with three stations"
  ) + 
  scale_colour_viridis_d()
  


weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = .3) +
  labs(
    title = "Temperature at three stations",
    x = "Mininmum daily temp (C)",
    y = "Maximum daily temp (C)",
    caption = "Data from rnoaa package with three stations"
  ) + 
  viridis::scale_colour_viridis(
    name = "location",
    discrete = TRUE ## Generate a discrete palette? (default: FALSE - generate continuous palette).
  )    
```


## Themes

```{r}
library(viridis)
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = .3) +
  labs(
    title = "Temperature at three stations",
    x = "Mininmum daily temp (C)",
    y = "Maximum daily temp (C)",
    caption = "Data from rnoaa package with three stations"
  ) + 
  scale_colour_viridis_d() +
  theme_bw() +   ##change every default things to black and white
  theme(legend.position = "bottom")  ## if we put this above `theme_bw`, it doesn't matters, bcs bw only changes default things!
```

you can also use `theme_minimal` (preferred), `theme_classic`, and even you could use `ggthemes::` package and try various themes contain in `ggthemes` package.

use `ggthemes` package ...
```{r}
library(viridis)
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = .3) +
  labs(
    title = "Temperature at three stations",
    x = "Mininmum daily temp (C)",
    y = "Maximum daily temp (C)",
    caption = "Data from rnoaa package with three stations"
  ) + 
  scale_colour_viridis_d() +
  ggthemes::theme_excel() +
  theme(legend.position = "bottom")
```
`theme_minimal()`
```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = .3) +
  labs(
    title = "Temperature at three stations",
    x = "Mininmum daily temp (C)",
    y = "Maximum daily temp (C)",
    caption = "Data from rnoaa package with three stations"
  ) + 
  scale_colour_viridis_d() +
  theme_minimal() +
  theme(legend.position = "bottom") ## if I want to do `theme_minimal` and change legend position both, I should put the minimal before legend, because minimal will COVER EVERYTHING before it!!
```


## Setting options

which is always put at the very beginning of our rmd file to setting things
```{r}
library(tidyverse)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```


## `data` in geoms (DATA ARRANGEMENT)

```{r}
central_park = 
weather_df %>% 
  filter(name == "CentralPark_NY")

waikiki = 
weather_df %>% 
  filter(name == "Waikiki_HA")

waikiki %>% 
  ggplot(aes(x = date, y = tmax, color = name)) +
  geom_point() +
  geom_line(data = central_park)  ## split a data set in two by adding a new graph
```

putting `data = ` in the `geom` function can using a different data set! 
If you just use `geom_line()` it will still work with `waikiki` data set.

## `patchwork`

```{r}
library(patchwork) ## load package

ggp_tmax_tmin = 
  weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = .3) +
  theme(legend.position = "none")

ggp_prcp_dens = 
  weather_df %>% 
  filter(prcp > 0) %>% 
  ggplot(aes(x = prcp, fill = name)) + 
  geom_density(alpha = .3) +
  theme(legend.position = "none")

ggp_tmax_date = 
  weather_df %>% 
  ggplot(aes(x = date, y = tmax, color = name)) +
  geom_point() +
  geom_smooth()

((ggp_tmax_tmin + ggp_prcp_dens) / ggp_tmax_date)
```


## Data Manipulation

quick example on factors

compare two plots, the second one rearrange the name according to the tmax value.

```{r}
weather_df %>% 
  ggplot(aes(x = name, y = tmax)) +
  geom_boxplot()

weather_df %>% 
  mutate(
    name = fct_reorder(name, tmax) ## arrange factor "name" in the order according to "tmax" value
  ) %>% 
  ggplot(aes(x = name, y = tmax)) +
  geom_boxplot()

weather_df %>% 
  mutate(
    name = factor(name), ## change location names from `chr` to `fct` variables.
    name = forcats::fct_relevel(name, c("Waikiki_HA")) ## will be reached out in the future
  ) %>% 
  ggplot(aes(x = name, y = tmax)) +
  geom_boxplot()
```

`fct_reorder()` function

What about tamx and tmin ...

```{r}
weather_df %>% 
  pivot_longer(
    tmax:tmin,
    names_to = "obs",
    values_to = "temperature"
  ) %>% 
  ggplot(aes(x = temperature, fill = obs)) +  ## `color = obs` is also ok
  geom_density(alpha = .3) +
  facet_grid(. ~ name)
```

```{r}
pulse_data = 
  haven::read_sas("./data/public_pulse_data.sas7bdat") %>%
  janitor::clean_names() %>%
  pivot_longer(
    bdi_score_bl:bdi_score_12m,
    names_to = "visit",
    values_to = "bdi",
    names_prefix = "bdi_score_"
  ) %>% 
  mutate(visit = recode(visit, "bl" = "00m"))
  
ggplot(pulse_data, aes(x = visit, y = bdi)) + 
  geom_boxplot()
```


in select function, you can use `start_with("pd_)` funtion to quickly select columns with same beginning!

This is a bonus practice!

First we manipulate our data!
```{r}
pup_data = 
  read_csv("./data/FAS_pups.csv", col_types = "ciiiii") %>%
  janitor::clean_names() %>%
  mutate(sex = recode(sex, `1` = "male", `2` = "female")) 

litter_data = 
  read_csv("./data/FAS_litters.csv", col_types = "ccddiiii") %>%
  janitor::clean_names() %>%
  separate(group, into = c("dose", "day_of_tx"), sep = 3)

fas_data = left_join(pup_data, litter_data, by = "litter_number") 
```
This step shows how to make many useful variables (i.e. the dose, days of treatment, outcomes...) shown clearly and exhaustively!
```{r}
fas_data %>% 
  select(dose, day_of_tx, starts_with("pd_")) %>% 
  pivot_longer(
    pd_ears:pd_walk,
    names_to = "outcome",
    values_to = "pn_day"
  ) %>% 
  drop_na() %>%
  mutate(outcome = forcats::fct_relevel(outcome, "pd_ears", "pd_pivot", "pd_walk", "pd_eyes")) %>% ## reorder the factors make it better
  ggplot(aes(x = dose, y = pn_day)) +  ## the y should be pn_days instead of outcome!
  geom_violin() +
  facet_grid(day_of_tx ~ outcome) ## we facet the day_of_tx against outcome!
```


```

